GMSE: an R package for generalised management strategy evaluation

### 2017-08-09 notebook/ms/ms.Rmd
Major restructuring of genetic algorithm section

Consistent with the MSE approach [@Bunnefeld2011], GMSE does not attempt to find optimal strategies or solutions for agents (stakeholders). Instead, genetic algorithms are used to find an adaptive strategy for each stakeholder in each time step.

A game-theoretic approach is compatible with agent-based modelling [@An2012; @Tesfatsion2017], but given the number of possible strategies available to agents, identifying optimal strategies is often intractable. Instead, genetic algorithms, which mimic the process of natural selection (mutation, recombination, selection, reproduction), are useful to find adaptive solutions for game strategies [e.g., @Balmann2000; @Tu2000; @Hamblin2013].

In keeping with the MSE approach [@Bunnefeld2011], GMSE does not attempt to find optimal strategies or solutions for stakeholders. Instead, within each time step of a simulation (Figure 1), a new genetic algorithm is called for each stakeholder to heuristically find strategies that are well-adapted to achieving their goals. Hence for the manager and each user, a unique population with a random set of possible values affecting policy or actions, respectively, is initialised. In each generation of the genetic algorithm, strategies crossover and mutate; when this results in strategies that are over-budget, values are iteratively decreased at random until budget constraints are satisfied. A fitness function then evaluates each strategy in the population, and a tournament is used to select the next generation of strategies [@Hamblin2013]. The genetic algorithm terminates when a minimum number of generations has passed and the increase in the fitness of the fittest strategy between the current and previous generation is below some threshold. The highest fitness strategy in the population then becomes the stakeholder's strategy in the current time step.

>>>>

Consistent with the MSE approach [@Bunnefeld2011], GMSE does not attempt to find optimal strategies or solutions for agents (stakeholders). Instead, genetic algorithms are used to heuristically find an adaptive strategy for each stakeholder in each time step. Critically, all stakeholders involved in resource conservation are constrained in their decision-making; managing and using resources takes effort (e.g., time or money), and effort expended in developing or enforcing one policy (for managers) or performing one action (for users) will be effort not expendable elsewhere [@Milner-Gulland2011]. In finding strategies, GMSE models this trade-off by setting a fixed budget for managers and users (Table 1). Allocations from a manager's budget can be used to increase the cost it takes a user to perform an action, and allocations from a user's budget can be used to perform the action at the cost set by the manager. Hence, stakeholders can have incomplete control over resource use and express competing management objectives.

In each new call of the genetic algorithm, a unique population of managers or users with random strategies is temporarily initialised. In each generation of the genetic algorithm, these strategies crossover and mutate; when this results in strategies that are over-budget, expenditures are iteratively decreased at random until budget constraints are satisfied. A fitness function then evaluates each strategy in the population, and a tournament is used to select the next generation of strategies [@Hamblin2013]. The genetic algorithm terminates when a minimum number of generations has passed and the increase in the fitness of the fittest strategy between the current and previous generation is below some threshold. The highest fitness strategy in the population then becomes the stakeholder's new strategy.
